### 2022-01-25 manuscript/manuscript.Rmd
Reword the methods section

The study protocol was approved by the local Ethics Committee (CES 192-14), and all participants were informed about the experiments purpose and protocol before giving written informed consent. A total of `r max(sample_size$pGRF_N)` adults were enrolled (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD). No participant reported any neurological or musculoskeletal limitations. Height and body mass were assessed following standard procedures with a stadiometer (model 213; Seca, Hamburg, Germany) and a digital scale (model 899; Seca, Hamburg, Germany), respectively.

After familiarization, participants performed a series of trials envolving jumps of different types and heights. First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments. Then, box jumps were performed from the floor to 5cm, 15cm and 30cm high boxes. Finally, continuous jumps were carried out in two different heights: 5cm and 15cm. Drop and box jumps were realized in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height. The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in force plates (1000Hz; AMTI Corporation, Watertown, USA). Participants were instructed to jump with no restrictions to arms movement, land two-footed and then remain stationary on the force plate for five seconds (except in the continuous jumps).

Throughout the protocol, three activity monitors (GT9X Link; 100Hz; ±16g range; ActiGraph, Pensacola, USA) were worn by each participant, at the following placements: i) at their right hip (along the anterior axillary line, at the level of the iliac crest); ii) at their lower back (at the midpoint between the two posterior superior iliac spines); and iii) at their right ankle (immediately superior to the lateral malleolus). Activity monitors at waist level were secured on the same elastic belt with a clip and the activity monitor placed on the ankle was secured by an elastic belt fixed with adhesive tape. In all placements, the accelerometer vertical axis was aligned with the standing body longitudinal axis.

The motion sensors used incorporate primary and secondary triaxial accelerometers. As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, to improve the data processing replicability only data from that accelerometer were used in this study. Data from FP and accelerometers were operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the *x*, *y* and *z* vectors. GRF was expressed in N and acceleration (ACC) in gravitational ACC units (1*g* = 9.807 m·s^-2^).

FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA) according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz). Then, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude. This cut-off frequency was selected based on visual inspection of the signals frequency spectrum after a Fast Fourier Transform (FFT). Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$). After that, both signals were adjusted using the time set by the systems clock and then synchronized by the maximum cross-correlation coefficient, with posterior manual adjustments if necessary. Next, peak ACC (pACC) were determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps. Peak GRF (pGRF) were then defined as the highest value between before and after 0.2 or 4 seconds, according to the jump type, after each pACC. The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1). The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).

@@

Statistical analyses were conducted using R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria). All statistical analyses were registered in an open platform, where R code utilized can be assessed (LINK). The statistically significant value was set as $\alpha$ = 0.05.

Linear mixed models were used to develop pGRF and pLR prediction equations for each vector (resultant and vertical) and accelerometer placement (ankle, lower back and  hip). Body mass was tested as predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models. All predictors were tested as fixed effects and were shown to be significant. Subject and the interaction between jump type and height were tested as random effects, and both factors have shown to improve the models. Final models were chosen according to the -2 log-likelihood statistic [@Field2012]. Traditional coefficient of determination (R^2^) was represented by conditional R^2^, which estimates the variance explained by the whole model [@Nakagawa2013].

Model validation was performed using the leave-one-out cross-validation [@Staudenmayer2012]. In this approach, each participant’s data is split into a testing dataset (one participant at a time), and the remaining data is in the training dataset. New linear mixed models with the same specification as determined for the whole sample are developed using the training dataset and then used to predict the outcome for the participant in the testing dataset. This process is, then, repeated for all participants (`r max(sample_size$pGRF_N)` times). Data from the testing dataset was used in the remaining statistical analyses.

>>>>

A total of `r max(sample_size$pGRF_N)` adults were enrolled in the study (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD). The protocol was approved by the local Ethics Committee (CES 192-14), and all participants were informed about the experiments purpose and protocol before giving written informed consent. No participant reported any musculoskeletal limitations. The measurement of height (stadiometer model 213; Seca, Hamburg, Germany) and body mass (digital scale model 899; Seca, Hamburg, Germany) was accomplished according to standard practice.

After familiarization, participants performed a series of trials envolving jumps of different types and heights. First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments. Then, box jumps were performed from the floor to 5cm, 15cm and 30cm high boxes. Finally, continuous jumps were carried out in two different heights: 5cm and 15cm. Drop and box jumps were executed in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height. The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in force plates (1000Hz; AMTI Corporation, Watertown, USA). Participants were instructed to jump with no restrictions to arms movement, land two-footed and then remain stationary on the force plate for five seconds (except in the continuous jumps).

Throughout the protocol, three activity monitors (GT9X Link; 100Hz; ±16g range; ActiGraph, Pensacola, USA) were worn by each participant, at the following placements: i) at their right hip (along the anterior axillary line, at the level of the iliac crest); ii) at their lower back (at the midpoint between the two posterior superior iliac spines); and iii) at their right ankle (immediately superior to the lateral malleolus). Monitors worn at hip and lower back were positioned on the same elastic belt with clips, and the ankle-worn monitor was fixed by an elastic belt and adhesive tape. Activity monitors were always positioned to ensure the alignment of the accelerometer vertical axis and the standing body craniocaudal axis.

The activity monitors used include a primary and a secondary triaxial accelerometers. As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, only data from this accelerometer were used in this study, to improve data processing replicability. Data from FP and accelerometers were operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the *x*, *y* and *z* vectors. GRF was expressed in N and acceleration (ACC) in gravitational ACC units (1*g* = 9.807 m·s^-2^).

The MATLAB software (Version 2019a, Mathworks, Natick, USA) was used to process data from both FP and accelerometers. Data were processed according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz). Then, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude. This cut-off frequency was selected based on visual inspection of the signals frequency spectrum after a Fast Fourier Transform (FFT). Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$). After that, both signals were adjusted using the time set by the systems clock and then synchronized by the maximum cross-correlation coefficient, with posterior manual adjustments if necessary. Next, peak ACC (pACC) were determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps. Peak GRF (pGRF) were then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type. The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1). The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).

@@

Statistical analyses were performed through the R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria). All code used to generate the results can be found in an open platform (https://github.com/verasls/jumping_GRF_ACC)(ref). Statistical significance was set as $\alpha$ = 0.05.

Linear mixed models were used to develop pGRF and pLR prediction equations for each vector (resultant and vertical) and accelerometer placement (ankle, lower back and  hip). Body mass was tested as predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models. These predictors were entered as fixed effects and kept in the model, as all have shown to be significant (*p* < 0.05). Subject and the interaction between jump type and height were tested as random effects, and both factors have shown to improve the models. Final models were chosen according to the -2 log-likelihood statistic [@Field2012]. The conditional coefficient of determination (R^2^), which estimates the variance explained by both fixed and random factors [@Nakagawa2013], was computed.

Model validation was performed using the leave-one-out cross-validation [@Staudenmayer2012]. This method consists in separating data from one of the subjects into a testing dataset, and keeping the remainder in the training dataset. Then, a new model is developed using the training dataset, with the same parameters that were defined using the whole sample. This new model is then applied to predict the outcome for the subject in the testing dataset. The whole process is repeated once for each participant (`r max(sample_size$pGRF_N)` times). Data from the testing dataset were used in all subsequent analyses.
